---
title: 'BM02 Annotation'
author: |
  | FHIL
  | Derrik Gratz
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
    css: "../config/robobook.css"
---

```{r, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE,
                      fig.width = 8, fig.height = 6,
                      cache=FALSE, cache.lazy = FALSE) 
```

This report is intended as an in-depth look at the decision points, intermediate data, and possible points of interest. 
It does not represent a final state of the data, and not every figure here is intended to be used for publication. 
Aesthetics and image quality can be improved for final versions of images.

# Natural clustering {.tabset}

**Non-optimized** clustering for each sample. This is clustering at a default resolution, intended to just show any
underlying community structure in the samples. After the annotation, I will perform a more currated clustering.

```{r, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
  chunktext <- '\n## {{kit}} {.tabset .unlist}\n'
  out = c(out, knit_expand(text = chunktext))
  for (sample in metadata$Sample[metadata$Kit == kit]) {
    ind <- metadata$Individual[metadata$Sample==sample]
    rep <- metadata$Replicate[metadata$Sample==sample]
    chunktext <- '\n### {{ind}}{{rep}} {.unlist}
      \n```{r, fig.width=6, fig.height=5}\n
      DimPlot(objs[["{{sample}}"]])
      \n```\n
      '
    out = c(out, knit_expand(text = chunktext))
  }
}

cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```

# Reference mapping {.tabset}

Expression matrices from published annotated datasets are used to predict celltype labels in our data. [SingleR](https://www.nature.com/articles/s41590-018-0276-y) is used as the prediction algorithm with data
from the [celldex](https://github.com/SingleR-inc/celldex) collection. The following datasets were used:

- Human primary cell atlas [Mabbott et al. 2013](https://doi.org/10.1186/1471-2164-14-632)
- Bulk sorted hematopoietic cells [Novershtern et al. 2011](https://doi.org/10.1016/j.cell.2011.01.004)
- Bulk sorted immune cells [Monaco et al. 2019](https://doi.org/10.1016/j.celrep.2019.01.041)

I also used Seurat's reference integration method with the following dataset:

- Single-cell PBMCs [Stoeckius et al. 2017](https://doi.org/10.1038/nmeth.4380)


```{r, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
  chunktext <- '\n## {{kit}} {.tabset .unlist}\n'
  out = c(out, knit_expand(text = chunktext))
  for (sample in metadata$Sample[metadata$Kit == kit]) {
    ind <- metadata$Individual[metadata$Sample==sample]
    rep <- metadata$Replicate[metadata$Sample==sample]
    chunktext <- "\n### {{ind}}{{rep}} {.unlist}
      \n```{r, fig.width=12, fig.height=10}\n
      grid.arrange(DimPlot(objs[['{{sample}}']], group.by = 'predicted.pbmc3k_seurat_annotations', label=TRUE),
        DimPlot(objs[['{{sample}}']], group.by = 'Novershtern.main.labels', label=TRUE),
        DimPlot(objs[['{{sample}}']], group.by = 'Mona.main.labels', label=TRUE),
        DimPlot(objs[['{{sample}}']], group.by = 'HPCA.main.labels', label=TRUE), nrow=2)
      \n```\n
      "
      out = c(out, knit_expand(text = chunktext))
  }
}


cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```

# Gene signatures {.tabset}

Seeing how celltype gene markers from literature correspond to calls from each method/reference.

```{r, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
  chunktext <- '\n## {{kit}} {.tabset .unlist}\n'
  out = c(out, knit_expand(text = chunktext))
  for (sample in metadata$Sample[metadata$Kit == kit]) {
    ind <- metadata$Individual[metadata$Sample==sample]
    rep <- metadata$Replicate[metadata$Sample==sample]
    chunktext <- '\n Sample {{ind}}{{rep}} \n'
    out = c(out, knit_expand(text = chunktext))
    for (ref in c('nova', 'hpca', 'mona', 'pbmc3k')){
      chunktext <- '\n### {{ref}} {.unlist}
      \n```{r, fig.width=40, fig.height=11}\n
      figures[[paste0("marker_dotplots_", "{{ref}}")]][["{{sample}}"]]
      \n```\n
      '
      out = c(out, knit_expand(text = chunktext))
    }
    
    ## only doing one sample
    break
  }
}

cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```
