---
title: 'BM02 Clustering'
author: |
  | FHIL
  | Derrik Gratz
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
    css: "../config/robobook.css"
---

```{r, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE,
                      fig.width = 8, fig.height = 6,
                      cache=FALSE, cache.lazy = FALSE) 
```

This report is intended as an in-depth look at the decision points, intermediate data, and possible points of interest. 
It does not represent a final state of the data, and not every figure here is intended to be used for publication. 
Aesthetics and image quality can be improved for final versions of images.

# Natural clustering {.tabset}

Clustering resolution for each sample was selected by optimizing for the silhouette score for a span of resolution values that
produced between 10 and 30 clusters.The Leiden algorithm implemented by iGraph was used for clustering.

```{r, eval=TRUE, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
  chunktext <- '\n## {{kit}} {.tabset .unlist}\n'
  out = c(out, knit_expand(text = chunktext))
  for (sample in metadata$Sample[metadata$Kit == kit]) {
    ind <- metadata$Individual[metadata$Sample==sample]
    rep <- metadata$Replicate[metadata$Sample==sample]
    chunktext <- '\n### {{ind}}{{rep}} {.unlist}
    \n```{r, fig.width=6.5, fig.height=5}\n
    DimPlot(objs[["{{sample}}"]], raster = TRUE, raster.dpi = c(512, 512))
    \n```\n
    '
    out = c(out, knit_expand(text = chunktext))
  }
}

cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```

# Gene signatures {.tabset}

Celltype gene markers from literature. **If you have more, please let me know so we can expand the database**. 

```{r, eval=TRUE, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
  chunktext <- '\n## {{kit}} {.tabset .unlist}\n'
  out = c(out, knit_expand(text = chunktext))
  for (sample in metadata$Sample[metadata$Kit == kit]) {
    ind <- metadata$Individual[metadata$Sample==sample]
    rep <- metadata$Replicate[metadata$Sample==sample]
    chunktext <- '\n### {{ind}}{{rep}} {.unlist}
    \n```{r, fig.width=40, fig.height=11}\n
    figures$marker_dotplots[[paste0("{{sample}}")]]
    \n```\n
    '
    out = c(out, knit_expand(text = chunktext))
  }
}

cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```

# Reference mapping {.tabset}

Summarizing top annotations from all references used (see previous report for more details)

```{r, eval=TRUE, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
  chunktext <- '\n## {{kit}} {.tabset .unlist}\n'
  out = c(out, knit_expand(text = chunktext))
  for (sample in metadata$Sample[metadata$Kit == kit]) {
    ind <- metadata$Individual[metadata$Sample==sample]
    rep <- metadata$Replicate[metadata$Sample==sample]
    chunktext <- '\n### {{ind}}{{rep}} {.unlist}
      \n```{r, fig.width=12, fig.height=7}\n
      figures[["annotation_DT"]][["{{sample}}"]]
      \n```\n
      '
      out = c(out, knit_expand(text = chunktext))
  }
}


cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```

# Final annotations {.tabset}

My first pass at cell annotation, manually aggregating calls from the reference sets and marker genes.

**This may not be the final annotations**

```{r, eval=TRUE, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
  chunktext <- '\n## {{kit}} {.tabset .unlist}\n'
  out = c(out, knit_expand(text = chunktext))
  for (sample in metadata$Sample[metadata$Kit == kit]) {
    ind <- metadata$Individual[metadata$Sample==sample]
    rep <- metadata$Replicate[metadata$Sample==sample]
    chunktext <- '\n### {{ind}}{{rep}} {.unlist}
      \n```{r, fig.width=13, fig.height=5}\n
      grid.arrange(
        DimPlot(objs[["{{sample}}"]], group.by = "annotations", raster = TRUE, raster.dpi = c(512, 512), label=TRUE),
        DimPlot(objs[["{{sample}}"]], group.by = "cell_labels.coarse", raster = TRUE, raster.dpi = c(512, 512), label=TRUE),
        nrow=1
      )
      objs[["{{sample}}"]]@meta.data$cell_labels.coarse |> 
        tabyl() |> 
        arrange(desc(n)) |>
        adorn_rounding(digits=2) |>
        knitr::kable(col.names=c("Cell label", "Count", "Percent"))
      \n```\n
      '
      out = c(out, knit_expand(text = chunktext))
  }
}
cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```
