---
title: 'BM Cost'
author: "FHIL\nDerrikGratz"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
    css: '../config/robobook.css'
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE,
                      fig.width = 8, fig.height = 6,
                      cache=FALSE, cache.lazy = FALSE) 
```

Trying some different representations of cost. Yes the color scheme is terrible right now. We can address that later if we choose to keep the kits separate like this.

# Kit index map

For comparisons with more variables, it's tough to organize all the data on the plot in a tidy way. For now, I've come up with this index mapping method, where the kits from a vendor are organized by their throughput and represented with a numeric index. These indices keep a consistent color across kits. The goal is to not bog down the figure with the details of the kit name, but they can refer back to this dictionary for the mapping info

```{r}
figures[['kit_index_color_table']]
```


***

# Per Kit

## Kit cost

The overall cost and value of a kit if it's entirely used.

```{r, fig.width=12, fig.height=7}
figures[['kit_cost']] 
```

Same thing with breaks, but this may be misleading

```{r, fig.width=12, fig.height=7}
figures[['kit_cost_breaks']] 
```

## Cost per cell

```{r, fig.width=12, fig.height=5}
figures[['cost_per_cell']] 
```

```{r, fig.width=12, fig.height=5}
figures[['price_per_cell_2']]
```

## Reagent cost table

```{r}
figures[['reagent_cost_table']]
```

***

# Per sample

## Cost per sample

Now the cost per sample. Note the different x axis scale. While this is a little misleading, some kits only have <4 samples, so they get squished and hard to see when stretched to the 96 samples to match parse. 

```{r, fig.width=12, fig.height=5}
figures[['cost_per_sample']]
```

## Total cost for N samples

A flipped variant of the cost per sample: the total cost for N samples. 

```{r, fig.width=12, fig.height=5}
figures[['running_cost']] 
```

## Minimum cost

The minimum cost of running N samples. This helps explain why the cost per sample is lower for the kits with more reactions. E.g. if you want to run 5-8 samples on flex and want to spend the minumum amount right now (ignoring that it might be cheaper 'per sample' to get a higher throughput kit), you'd buy two 4x1 kits rather than the 4x4 kit. 

```{r, fig.width=12, fig.height=5}
figures[['n_samples_min_cost']]
```

## Cells per sample

```{r, fig.width=12, fig.height=5}
figures[['cells_per_sample']]
```

***

# Sequencing cost

## Example table

```{r}
figures[['seq_cost_table']]
```


## Seq cost per depth

```{r}
figures[['seq_cost_curves']]
```

***

# Example experiment cost table

```{r}
figures[['cost_table']]
```

