---
title: 'BM02 Data filtering'
author: "FHIL\nDerrikGratz"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
    css: '../../../config/robobook.css'
---

```{r, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE,
                      fig.width = 8, fig.height = 6,
                      cache=FALSE, cache.lazy = FALSE) 
```

# Filtering thresholds

Filtering cells by consistent global thresholds.

```{r}
stack(global_cutoffs) %>%
  separate(ind, c('metric', 'direction'), '\\.') %>%
  dcast(metric~direction, value.var = 'values') %>%
  knitr::kable(digits=2)
```

# Filtering Metrics

Looking at individual metrics and how many cells would be removed when filtering by global thresholds.

## nFeature

:::: {class='fluid-row'}

::: {class='col-md-6'}

```{r, fig.height=9, fig.width=4}
figures[['nFeature_vln_facet']]
```

:::

::: {class='col-md-6'}

```{r, fig.height=9, fig.width=6}
figures[['filtering_likert_nFeature_RNA']]
```

:::

::::

## nCount

:::: {class='fluid-row'}

::: {class='col-md-6'}

```{r, fig.height=9, fig.width=4}
figures[['nCount_vln_facet']]
```

:::

::: {class='col-md-6'}

```{r, fig.height=9, fig.width=6}
figures[['filtering_likert_nCount_RNA']]
```

:::

::::

## mito

:::: {class='fluid-row'}

::: {class='col-md-6'}

```{r, fig.height=9, fig.width=4}
figures[['mtRatio_vln_facet']]
```

:::

::: {class='col-md-6'}

```{r, fig.height=9, fig.width=6}
figures[['filtering_likert_mtRatio']]
```

:::

::::

## ribo

:::: {class='fluid-row'}

::: {class='col-md-6'}

```{r, fig.height=9, fig.width=4}
figures[['rbRatio_vln_facet']]
```

:::

::: {class='col-md-6'}

```{r, fig.height=9, fig.width=6}
figures[['filtering_likert_rbRatio']]
```

:::

::::

# Doublets

Doublets called by the `scDblFinder` algorithm.

```{r}
figures[['doublet_portion_boxchart']]
```

# Filtering overlap {.tabset}

Of the possible filtering metrics, is there some redundancy? 
Are some thresholds dominating the filtering?
And are doublets sufficiently removed by threshold filtering? 
It looks like `scDblFinder` is labeling many additional cells beyond those with 
high UMI/gene counts, likely due to heterotypic expression profiles. 
The doublet finding algorithm could also be viewed as too conservative or lacking accuracy & eliminating real data. 

```{r}
chunk_text <- '
\n## {{kit}} {.unlist}
\n\n```{r, fig.width=14, fig.height=4.5}
samples <- metadata$Sample[metadata$Kit == "{{kit}}"]
gridExtra::grid.arrange(grobs = euler_plots[samples], nrow = 1)
\n```\n'
```

```{r, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
  out = c(out, knit_expand(text = chunk_text))
}

cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```

# Cell recovery {.tabset}

What portion of the capture is removed by QC filtering, identified as doublets, etc.

## Counts

```{r, fig.width=12}
figures[['cell_recovery_count_barchart']]
```

## Proportions

```{r, fig.width=12}
figures[['cell_recovery_portion_barchart']]
```
