---
title: 'BM02 celltype resolution'
author: |
  | FHIL
  | Derrik Gratz
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
    css: '`r here("config/robobook.css")`'
---

```{r, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE,
                      fig.width = 8, fig.height = 6,
                      cache=FALSE, cache.lazy = FALSE) 
```

The goal is to see how strongly a celltype can be detected in the sample/kit.

# Annotation score average

Showing average annotation score for each cell class in the pbmcsca reference.
This could be a little biased as many of the cells in this reference are from 
10x kits. 

Cell counts by method in reference:

```{r}
pbmcsca$Method |> table()
```

For each celltype manually annotated in the datasets, how strong were the 
prediction scores for that celltype in those cells?

```{r}
figures[['annotation_scores_kit_fine']]
```

# Annotation confusion matrix {.tabset}

For all celltype labels, how specific are the annotations? Is it harder to resolve
some celltypes from others using annotation mapping?

```{r, eval=TRUE, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
  chunktext <- '\n## {{kit}} {.unlist}\n
  \n```{r, fig.width=11, fig.height=9}\n
  figures[["annotation_confusion_mat_fine"]][["{{kit}}"]]
  \n```\n
  '
  out = c(out, knit_expand(text = chunktext))
}

cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```

# Module definitions

To avoid potential bias with the reference dataset being largely 10x or droplet-based
samples, we can try to use 'objective' gene lists. However, there is no universally
agreed upon lists. 

Here, I've aggregated marker genes retrieved from literature, queried against [cellxgene](https://cellxgene.cziscience.com/gene-expression), and present
in 10+ cells of 2+ samples in all kits (besides Flex). 

**If you have more, please let me know so we can expand the database**. 

```{r}
# celltype_markers
celltype_markers_modules
```

# Gene signatures {.tabset}

Details for expression of every gene in the module for each sample

```{r, eval=TRUE, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
  # chunktext <- '\n## {{kit}} {.tabset .unlist}\n'
  # out = c(out, knit_expand(text = chunktext))
  # for (sample in metadata$Sample[metadata$Kit == kit]) {
  #   ind <- metadata$Individual[metadata$Sample==sample]
  #   rep <- metadata$Replicate[metadata$Sample==sample]
    chunktext <- '\n## {{kit}} {.tabset .unlist}\n
    \n```{r, fig.width=27, fig.height=6}\n
    figures$marker_dotplots[[paste0("{{kit}}")]]
    \n```\n
    '
    out = c(out, knit_expand(text = chunktext))
  # }
}

cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```

# Module Dotplots

Module scoring performed by [UCell](https://doi.org/10.1016/j.csbj.2021.06.043).

Within celltypes from each kit, how did they score with the modules? 
Comparing the ability to distinguish celltypes across kits.

Using a shared set of genes present in every kit (including Flex).
<!-- ## Coarse -->

<!-- ```{r} -->
<!-- figures[['module_scores_kit_coarse']] -->
<!-- ``` -->

<!-- ## Fine -->

```{r}
figures[['module_scores_kit_fine']]
```

# Confusion matrices {.tabset}

For all celltype labels and modules, how specific are the modules? Is it harder to resolve
some celltypes from others?

<!-- ## Coarse {.tabset} -->

<!-- ```{r, eval=TRUE, results='asis'} -->
<!-- out <- NULL -->
<!-- for (kit in unique(metadata$Kit)) { -->
<!--   chunktext <- '\n### {{kit}} {.unlist}\n -->
<!--   \n```{r, fig.width=12, fig.height=11}\n -->
<!--   figures[["confusion_mat_coarse"]][["{{kit}}"]] -->
<!--   \n```\n -->
<!--   ' -->
<!--   out = c(out, knit_expand(text = chunktext)) -->
<!-- } -->

<!-- cat(knit_child(text = out, quiet = TRUE), sep = '\n') -->
<!-- ``` -->

<!-- ## Fine {.tabset} -->

```{r, eval=TRUE, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
  chunktext <- '\n## {{kit}} {.unlist}\n
  \n```{r, fig.width=11, fig.height=9}\n
  figures[["confusion_mat_fine"]][["{{kit}}"]]
  \n```\n
  '
  out = c(out, knit_expand(text = chunktext))
}

cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```
