---
title: 'BM 3P WT DGE'
author: |
  | FHIL
  | Derrik Gratz
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
    css: '../../config/robobook.css'
---

```{r, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE,
                      fig.width = 8, fig.height = 6,
                      cache=FALSE, cache.lazy = FALSE) 
```

# PCA

## All kits

PCA of the top 2000 variable genes in all data.

```{r}
figures[['PCA_of_kits']]
```

## By kit {.tabset}

PCA of the top 2000 variable genes per kit.

```{r, eval=TRUE, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
    chunktext <- '\n### {{kit}} {.unlist}\n
    \n```{r, fig.width=7.5, fig.height=6}\n
    figures[[paste0("PCA_", "{{kit}}")]]
    \n```\n
    '
    out = c(out, knit_expand(text = chunktext))
}

cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```

# Corr mat {.tabset}

Spearman correlation of the top 200 variable genes per celltype within kit. This is a measure of replicability across samples. Higher correlation indicates top genes are similarly ranked. 

First, the average of the two replicate pairs is shown here

```{r, fig.width=8, fig.height=8}
figures[['rep_corr_summary']]
```

Then a breakdown of all kits (this may be excessive)

```{r, eval=TRUE, results='asis'}
out <- NULL
for (kit in unique(metadata$Kit)) {
    chunktext <- '\n## {{kit}} {.unlist}\n
    \n```{r, fig.width=8, fig.height=5}\n
    figures[[paste0("corr_mat_pearson", "{{kit}}")]]
    \n```\n
    '
    out = c(out, knit_expand(text = chunktext))
}

cat(knit_child(text = out, quiet = TRUE), sep = '\n')
```

