---
title: 'BM01 TCR rep analysis'
author: "FHIL\nDerrikGratz"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
    css: '../../../config/robobook.css'
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.width = 8, fig.height = 6,
                      cache=FALSE, cache.lazy = FALSE) 
```

# Clone recovery {.tabset}

<!-- ## Reported -->

<!-- As reported by the pipelines. The TRA and TRB columns also include cells that had BOTH TRA/TRB, so they  -->
<!-- reflect the total % of cells with that chain regardless of if the other was present.  -->

<!-- ```{r} -->
<!-- figures[['reported chain recovery']] -->
<!-- ``` -->


<!-- ## Calculated {.tabset} -->

Calculated from the final output files. This has a more granular breakdown of chain recovery, distinguishing
cells with multiple viable chains for either TRA or TRB

## counts

Raw counts

```{r}
figures[['chain recovery count']]
```

## proportions {.active}

```{r}
figures[['chain recovery % recovered']]
```

```{r}
figures[['chain recovery stacked']]
```

***

# Unique clonotype recovery

```{r}
figures[['unique clone recovery count']]
```

***

# Diversity

A higher inverse Simpson index value indicates more unique members and more even distribution of presence.

## All clones

```{r}
figures[['simpson index']]
```

## Clean clones

TRA:TRB perfect pairs only. No multiple or missing chains.

```{r}
figures[['simpson index perfect clones only']]
```

