---
title: 'BM01 TCR rep analysis'
author: "FHIL\nDerrikGratz"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
    css: '../../../config/robobook.css'
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.width = 8, fig.height = 6,
                      cache=FALSE, cache.lazy = FALSE) 
```

# clone recovery {.tabset}

## Reported

As reported by the pipelines. The TRA and TRB columns also include cells that had BOTH TRA/TRB, so they 
reflect the total % of cells with that chain regardless of if the other was present. 

```{r}
figures[['reported chain recovery']]
```


## Calculated {.tabset}

Calculated from the final output files. This has a more granular breakdown of chain recovery, distinguishing
cells with multiple viable chains for either TRA or TRB

### counts

Raw counts

```{r}
figures[['chain recovery count']]
```

### proportions {.tabset}

#### All T Cells

```{r}
figures[['chain recovery % all']]
```

#### Productive T cells

```{r}
figures[['chain recovery % recovered']]
```

# {- .unlisted}

***


# Diversity

A higher inverse Simpson index value indicates more unique members and more even distribution of presence.

```{r}
figures[['simpson index']]
```

